---
title: "Data Cleaning Documentation"
author: "Aud Halbritter"
date: "18 Juli 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
#knitr::read_chunk("Climate Data Check.R")

library("tidyverse")
library("lubridate")
library("knitr")
library("ggplot2")

load("traits/traits.Rdata")


```

## Content

This document contains explanations for the collection and processing of the following data sets:
* Plant community composition
* Leaf functional traits
* Biomass

## Site information
The study is conducted along an elevational gradient ranging from sub-alpine to alpine meadows in the Hengduan Mountain of southwest China. The elevational gradient contains of four sites: lowland (L 3000m), middle (M 3500m), alpine (A 3850m) and high alpine (H 4100m).

Coordinates?

## Experiments
Each site is divded into 5 (H) - 7 blocks within a fence to protect the site from grazing. Each block contains up to 7 different plots: OTC (open top chamber), C (untouched control), 0 (local transplant), 1 (transplant one step down, warm1), 2 (transplant one step up, cold1), 3 (transplant all the way down, warm3), 4 (transplant all the way up, cold4).

I think it is 0 in the trait data set and O in the community data set. Should be the same!


# Plant community composition

#### Data set and collection
The dataset contains plant community composition data from the an elevational gradient in the Hengduan Mountains

The dataset contains the sub-plot frequency, cover, and some performance data on 108 taxa of vascular plants, along with some plot-level predictor variables, in 140 25 x 25 cm vegetation plots.

The dataset covers the years 2012-2015. 2016 has not been added yet.


For each species in each subplot it was noted down whether it was present as sterile (x), dominant (covered more than 50% of the sub-plot: D), fertile (circle around the X or D), juvenile (J), or a seedling (S) ??? juvenile and seedlings probably not? 

For the entire 25 x 25 cm plot, the percentage cover of each vascular plant species was then visually estimated. 

Other variables that were measured for each plot:
- vegetation height (mean of 4 measurements)
- moss % cover
- moss layer depth (mean of 4 measurements)
- litter % cover 
- bare ground % cover
- rock % cover
- date of analysis
- recorder / scribe (if any)
- comments

Other variables that were measured for each sub-plot:
- moss % cover
- litter % cover 
- bare % ground cover
- rock % cover

Nomenclature follows the Plant list???
Each species of vascular plant was classified by functional type (graminoid, forb, woody)

#### Data cleaning
The information from the field data sheets was manually entered into digital worksheets, manually proof-read and stored in the a database. 
Species’ occurrences were compared over the years to search for and correct data identification, recording and entry errors. Specifically, maps of each species in each sub-turf over the years were scrutinized to search for species identification errors. Based on this, corrections were made. for example:
If it was obvious from the data that a species of a difficult taxa, such as Carex, had occasionally been misidentified in a year when it did not flower, the erroneous IDs were corrected.   
If it was obvious that a  species could not be accurately identified to the species level on the spatial and temporal scale of the dataset, e.g., because they were not identifiable even to experts when sterile, the species were merged.
All corrections are  recorded and documented. All post-proofreading corrections (e.g., taxonomic mergers) are done in code post-database.


## Leaf functional traits

#### Data set and collection
Functional traits include:
LDMC (Leaf Dry Matter Content)
SLA (Specific leaf area)
LA (Leaf area)
LT (Leaf thickness)
LN (Leaf Nitrogen)
LC (Leaf Carbon)

##### Gradient leaves
The dataset contains leaf functional trait data from plant species at 4 sites along an elevational gradient.The data was collected in 2015 and 2016.

##### Leaf sampling
Up to 5 leaves from preferably different healthy individuals were collected outside the experimental sites. 

Singular leaves were then separated from the shoot or the individual for the measurements. Collected leaves from each plot were put in zip-lock plastic bag, per species, and stored in a cooler to keep them from drying. The leaves were scanned, weighing (wet and dry), and leaf thickness was measured. All measurements were completed within 24 hours of collection.



##### Difference in data collection in 2015
* Leaves were all collected on one day. All species that could be found were collected, but not necessarily always healthy plants.
* Not always sure if individuals are different


##### Experimental leaves
The dataset contains leaf functional trait data from common plant species in plots from an elevational transplant and climate warming experiment. Trait data of the most common species, based on the community data of the Transplant project and excluding graminoids, were collected. All data was collected in summer 2016. 

Used vegetational data from previous year to find the most abundant species at the four locations. The selected species is as followed: 
Aletris pauciflora, Allium prattii, Artemesia flaccida, Cerastium szechuense, Epilobium fangii, Galium hoffmeisteri, Hypericum wightianum, Pedicularis davidii, Plantago asiatica, Polygonum macrophyllum, Polygonum viviparum, Potentilla leuconota, Pseudostellaria davidii, Ranunculus tanguticus, Veronica szechuanica, Viola biflora var. rockiana. 

Following this species list, up to 5 leaves from preferably different individuals within each plot (25 x 25 cm) were collected. If the turf had only one individual of a target species, only one leaf was collected. When the individual did not look healthy, no leaves were collected. 
In the OTCs leaves were collected throughout the entire chamber. 
	In Control plots, leaves from around the edges could also be collected.
Whereas in all transplant treatments leaves were only collected from individuals definitely inside the plots. 




#### WET VS. DRY MASS
colour Tiny wet mass: Wet_Mass_g < 0.0005
```{r, echo = FALSE}
traits %>% mutate(year = as.factor(year(Date))) %>%
  ggplot(aes(x = Wet_Mass_g, y = Dry_Mass_g, colour = Wet_Mass_g < 0.0005)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```

colour: Wet < Dry

```{r, echo = FALSE}
traits %>% mutate(year = as.factor(year(Date))) %>%
  ggplot(aes(x = Wet_Mass_g, y = Dry_Mass_g, colour = Wet_Mass_g < Dry_Mass_g)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```

2016 - problems with area

```{r, echo = FALSE}
traits %>% mutate(year = as.factor(year(Date))) %>%
  filter(year == 2016) %>% 
  ggplot(aes(x = Wet_Mass_g, y = Dry_Mass_g, colour = AreaFlag)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```  

2016 - problems with dry mass

```{r, echo = FALSE}
traits %>% mutate(year = as.factor(year(Date))) %>%
  filter(year == 2016) %>% 
  ggplot(aes(x = Wet_Mass_g, y = Dry_Mass_g, colour = DryFlag)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```


## DRY VS. AREA 
Problems with Area

```{r, echo = FALSE}
traits %>% mutate(year = as.factor(year(Date))) %>%
  ggplot(aes(x = Dry_Mass_g, y = Leaf_Area_cm2)) + 
  geom_point() +   
  geom_abline(intercept = 0, slope = 1, colour = "red") +
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```

## SLA
SLA > 500

```{r, echo = FALSE}
traits %>% 
  mutate(year = year(Date)) %>% 
  arrange(SLA_cm2_g) %>%
  ggplot(aes(x = Leaf_Area_cm2, y = Dry_Mass_g, colour = SLA_cm2_g > 500)) + 
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) + 
  scale_x_log10() + 
  scale_y_log10() + 
  facet_wrap(~ year)
```


## Biomass
#### Data set and collection
The biomass data set contains 20 x 50x50cm plots from each site. All biomass was harvested from each plot and sorted to species and weighted. The cover for each species has been estimated and height of up to 10 individuals has been measured.

#### Problems with the data set
* There are still duplicate plants in the data set that should be merged
* it is unclear how the height was measured. How were the 1 random leaves selected. Needs clarification from China
* How good is the cover estimate?
* Can we trust the graminoid species or should these be merged?
* There are 2 plots that contain > 20% seedlings (impossible!)
* a couple of seedlings are huge


