---
title: "Quality control checks on transplant community data"
author: "Richard J. Telford"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
classoption: a4paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

#Import data
```{r, importChunk}
#load packages
library("tidyr")
library("plyr")
library("DBI")# also needs RMySQL installed
library("dplyr")
library("ggplot2")

#source functions
source("R/load_subturfcomm.R")
source("R/load_comm.R")

#make database connection
con <- dbConnect(RMySQL::MySQL(), group = "transplant")

#load cover data and metadata
subturf_thin <- load_subturfcomm(con = con)
#  spread(subturf.thin, key = "subTurf", value = "adult", fill = 0)
cover_thin <-load_comm(con = con)
```

#Missing turfs

The following turfs have data missing from at least one year. I know that some of these are yak damaged - it would be good to have this documented somewhere - but others should have data

```{r}
##missing turfs
missingturfs <-  subturf_thin %>% 
  group_by(turfID, year, originSiteID, destSiteID, TTtreat) %>%
  summarise(n = "") %>% 
  spread(key = year, value = n, fill = "miss") %>%
  ungroup() %>%
  filter(drop = rowSums(. == "miss") > 0) %>%
  select(-originSiteID)

knitr::kable(missingturfs)
```

#Missing species

There are some species that are present in the subplot data that are missing from the cover data and vice-versa.

```{r}
#check subturf and cover contain same species
subturf_freq <- subturf_thin %>% 
  group_by(turfID, species, year) %>%
  summarise(n = n())

#merge cover and subturf
mergedCoverFreq <- merge(cover_thin, subturf_freq, all = TRUE)
```

The following species occurences are in subplots, but not in cover (n gives number of subplots)
```{r}
inSubplot <- mergedCoverFreq %>% filter(is.na(cover)) %>% select(turfID, year, species, n)
knitr::kable(inSubplot)
```

The following species are in the cover data, but not in the suplots

```{r}
inCover <- mergedCoverFreq %>% filter(is.na(n)) %>% select(turfID, year, species, cover)
knitr::kable(inCover)
```

#Non-integer cover values

In seedclim, cover is estimated to integer percent. In the transplant data, there are many non-integer values.

```{r nonIntegerChunk}
nonInt <- cover_thin %>% 
  filter(cover != round(cover)) %>% 
  group_by(cover) %>% 
  summarise(n_occurrences = n())
knitr::kable(nonInt)
```

Many of these are probably correct, but some, such as, `r max(cover_thin$cover[cover_thin$cover != round(cover_thin$cover)])` may be data-entry errors.

These are the examples with a cover greater than 5%.
```{r}
nonInt2 <- cover_thin %>% filter(cover > 5 & cover != round(cover)) %>% select(turfID, species, cover)
knitr::kable(nonInt2)
```

Please can you check these, and initiate proof reading if required.

#Species with large changes in cover

Some species have large change in cover between years (here defined as > 25%). These may be real (some maybe because of the missing turfs), others maybe because of data input errors or field errors. You probably have an idea of which species can fluctuate in abundance and which are generally consistent, and then check probably problems.

So far, I have only checked the control, as large changes in cover can be expected in the treatment turfs, so these need checking by hand.

```{r}
changes <- cover_thin %>% 
  select(originSiteID, turfID, destSiteID, TTtreat, year, species, cover, speciesName) %>%
  filter(TTtreat %in% c("control", "local")) %>% 
  spread(key = year, value = cover, fill = 0) %>%
  gather(key = year, value = cover, -originSiteID, -turfID, -destSiteID, -TTtreat, -species, -speciesName) %>%
  group_by(turfID, species) %>%
  mutate(RANGE = diff(range(cover))) %>%
  spread(key = year, value = cover)

knitr::kable(
  changes %>% ungroup() %>% filter(RANGE >= 25) %>% arrange(turfID) %>% select(-originSiteID, -destSiteID, -RANGE, -species)
)
```

#Sum of cover patterns

Some of the sites have very high sum of covers in some years

```{r}
sumcover <- cover_thin %>%
  group_by(turfID, year, TTtreat, destSiteID, originSiteID) %>%
  summarise(sumCover = sum(cover))

ggplot(sumcover, aes(x = year, y = sumCover, colour = factor(turfID, levels = sample(unique(turfID))))) +
  geom_line(show.legend = FALSE) +
  scale_x_continuous(breaks = unique(cover_thin$year), labels = paste0("`", substr(unique(cover_thin$year), 3, 4))) +
  facet_grid(originSiteID ~ TTtreat) +
  labs(x = "Year", y = "Sum of Covers")

knitr::kable(
  sumcover %>% filter(sumCover >180) %>% select(-originSiteID)
)
```

#Richness patterns 

This figure shows changes in richness over time for the different sites and treatments.

```{r}
cover_thin %>% group_by(destSiteID, TTtreat, turfID, year) %>%
  summarise(richness = n()) %>%
  ggplot(aes(x = destSiteID, y = richness, fill = as.factor(year))) +
    geom_boxplot()+
    facet_wrap(~TTtreat) +
    labs(x = "Destination Site", y = "Turf Richness", fill = "Year")
```

There are some interesting patterns, but also some large changes in the controls (perhaps related to the fence effect) that we need to understand.

#
#Species with large changes in commonness

These are species that, in the control plots, have the largest (penalised) relative changes in the number of plots occupied.
```{r}
ro <- cover_thin %>%
  filter(TTtreat %in% c("local", "control")) %>%
  group_by(species, year) %>%
  summarise(n = n()) %>%
  spread(key = year, value = n, fill = 0) %>%
  ungroup() %>%
  mutate(max = apply(.[,-1], 1, max), min = apply(.[,-1], 1, min), variability = (max - min)/(max + 10)) %>%
  arrange(desc(variability)) %>%
  select(-max, -min, - variability)
knitr::kable(head(ro, n = 20))
```
