---
title: "Climate Data"
author: "Aud Halbritter"
date: "27 3 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
knitr::read_chunk("Climate Data Check.R")

library("tidyverse")
library("lubridate")
library("knitr")

load("weather_unclean.Rdata")
load("climate.Rdata")
load("climate_month.Rdata")

```


## Data cleaning #1

Units

```{r UncleanTair, echo = FALSE}

```

## Data cleaning #2

Spikes

```{r UncleanwaterContent5, echo = FALSE}

```



## Air temperature

Gradient: 2m; OTC: 30 cm

```{r DataTair, echo = FALSE}

```


## Ground temperature

```{r DataTsoil0, echo = FALSE}

```


## Soil temperature -5cm

Not much data from the gradient; increasing variation at L site

```{r DataTsoil5, echo = FALSE}

```


## Soil temperature -20cm

Data ok, but we do not expect much difference at -20cm

```{r DataTsoil20, echo = FALSE}

```


## Gradient
```{r AirTemperatureGradient, echo = FALSE}

```



## Gradient vs OTC

July tempereature

```{r GradientVSOtc, echo = FALSE}

```




## Soil moisture

```{r DatawaterContent5, echo = FALSE}

```



## Precipitation

No data

```{r Datarain, echo = FALSE}

```


## PAR

Problematic cut off

```{r DataPAR, echo = FALSE}

```


## Solar Radiation

Cut off

```{r DatasolarRadiation, echo = FALSE}

```


## Relative humidity

```{r DataRH, echo = FALSE}

```

