---
title: "cover > maximum possible cover"
author: "Richard J. Telford"
date: "November 16, 2016"
output: html_document
---

```{r importChunk, echo = FALSE, message = FALSE}
#load packages
library("tidyr")
library("DBI")# also needs RMySQL installed
library("dplyr")
library("ggplot2")

#source functions
source("R/load_subturfcomm.R")
source("R/load_comm.R")

#make database connection
con <- dbConnect(RSQLite::SQLite(), dbname = "data/transplant.sqlite")

#load cover data and metadata
subturf_thin <- load_subturfcomm(con = con)
#  spread(subturf.thin, key = "subTurf", value = "adult", fill = 0)
cover_thin <-load_comm(con = con)

#merge cover amd subturf

subturf_cover_thin <- left_join(subturf_thin, cover_thin)

```

Each of the 25 subturfs covers 4% of the turf.

The maximum possible cover possible is equal to the number of occupied subturfs times 4 %. Some species have a much higher cover, suggesting that either cover is over estimated or not all the subturfs have been counted.

Below I show species that have a cover greater than 1.25 times the maximum possible cover (this avoids reporting species with small discrepancies, for example cover = 5 in one subturf.)

```{r}
out <- subturf_cover_thin %>% 
  group_by(year, turfID, species, cover) %>% 
  summarise(maxPos = n() * 4) %>% 
  filter(cover > (maxPos * 1.25)) 
knitr::kable(out)
```
