---
title: "Transplant Climate Metadata"
author: "Richard J. Telford"
date: "September 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("ggplot2")
library("dplyr")
library("tidyr")

#load data
load("otc_month.Rdata")
load("month_annual2016-09-21.Rdata")
```

## Overview

There are climate loggers at the four TRANSPLANT sites, running from `r format(min(monthly$month), "%B %Y")`. Data are recorded (usually) every 10 minutes. Each site has sensors recording the following variables (units and variable name given in brackets)

 - Air temperature at (?) 2 m (°C, Tair).
 - Soil surface temperatures (°C, Tsoil0)
 - Soil temperatures at 5 and 20 cm depth (°C, Tsoil5, Tsoil20)
 - Relative humidity at (?) 2 m (%, RH)
 - Soil moisture at 0, 5 and 20 cm depth (m^3^/m^3^, waterContent0,waterContent5, waterContent20)
 - Precipitation (probably mm, rain)
 - Wind speed (corrupt, windSpeed)
 - Wind direction (°, windDirection)
 - Solar radiation (W/m², solarRadiation)
 - Photosynthetically active radiation (uE, PAR)"PAR""solarRadiation"    
 - UV radiation (umolm^-2^s, UV)
 
In addition, there are logger within the Open Top Chamber recording  

 - Air temperature at (?) 30 cm (°C, Tair30).
 - Soil surface temperatures (°C, Tsoil0)
 - Soil temperatures at 5 and 20 cm depth (°C, Tsoil5, Tsoil20)
 - Relative humidity at (?) 2 m (%, RH)
 - Soil moisture at 0, 5 and 20 cm depth (m^3^/m^3^, waterContent0,waterContent5, waterContent20)
 - Photosynthetically active radiation (uE, PAR)"PAR""solarRadiation"    

Logger and sensor make/model details are not currently available.

Monthly and annual mean data for each station are available.

Data are subject to change when issues described below are fixed. 

## Data processing

The data was provided in excel files, with the site and OTC data formatted differently.

Code for reading the data from the excel files is in files 'read.climate2.r' and 'OTC_climate.R'. These are archived versions of the code; the working version of the code in the TRANSPLANT github repository.

Essentially, both files search the specified data directory for excel files and then loop through the list of file, extracting the contents of each and combining them into a large data.frame. Various steps are then taken to remove unambiguiously erroneous data and duplicate measurements before calculating monthly and then annual means (sums for precipitation).

Data for a month are given only if there is at least three weeks worth of data (ie 3 * 7 * 24 * 6 observations). To calculate annual means/sums, means were first calculate for each month so that uneven amount of data for different times of the year did not bias the results.

The site climate data files were corrupt, so the code is far more complex than would otherwise have been the case. The OTC files were malformatted, so the fast `readxl` package could not be used, and the very slow `gdata::read.xls` had to be used instead.

Data cleaning removed impossible values (temperatures of 100 °C), data with implausibly low variance and data that were otherwise impossible. Much data was lost in this process, including

 - Precipitation at all sites except A (because the values were derived from windspeed)
 - Wind direction at site A (because the values were derived from windspeed)
 - Much soil temperature data (impossible values or implausible variance)
 - UV data from site H (spurious trend)

The remaining data still has some issues

 - temperature data has spikes. These will make minimal difference to monthly or annual means, but need to be cleaned before high resolution data can be used
 - some soil temperature data has changes to the high-frequency variance over time. This probably doesn't affect the monthly or annual means. 
 - Water content data are dubious. Some of the data look good, but other have spurious trends or lack characteristic features such as a drop in apparent mositure during winter. Offsets between the sensors at different depths and under the OTC may reflect poor calibration or real variability. Use these data with extreme caution.
 - Wind speed data have a mix of units, which probably largely determines the trend in windspeed. Until this problem is fixed, use these data with extreme caution (and ignore the trend).
 
## Importing the data

The data are provided in .Rdata files. These can be read into R with the `load` command, and ensures that dates etc are correct. The data are in a tidy format: one row - one observation. This code will load the data and packages needed to plot them.
 
```{r loadStuff, eval = FALSE} 
```

## Annual means
```{r}
knitr::kable(
  x = annual %>% spread(key = variable, value = value),
  digits = 1)
```


## Plotting the data

Monthly site data

```{r}
ggplot(monthly, aes(x = month, y = value, colour = site)) + geom_path() + geom_point(size = 0.1) + facet_wrap(~variable, scales = "free_y")
```

Monthly OTC data

```{r}
ggplot(otc_month, aes(x = month, y = value, colour = site)) + geom_path() + geom_point(size = 0.1) + facet_wrap(~variable, scales = "free_y")
```


##Comparison OTC with external weather


### Temperature

Air and surface temperatures are typically higher in the OTC. Note that the air temperatures are not (as I understand) properly comparable as the OTC air temperature sensor is at a different height to the site air temperature sensor. Soil temperatures are similar.

```{r}
all_month <- rbind(
  cbind(sensor = "OTC", otc_month),
  cbind(sensor = "Site", monthly)
)

g <- all_month %>% filter(variable == "Tsoil0") %>% 
  ggplot(aes(x = month, y = value, colour = sensor)) + 
    geom_path() + 
    geom_point(size = 0.1) +
    facet_wrap(~site)

g

g %+% (all_month %>% filter(variable == "Tsoil5"))
g %+% (all_month %>% filter(variable == "Tsoil20"))
g %+% (all_month %>% filter(variable == "Tair30" | variable == "Tair"))
```

## About

The file `ClimateMeta.html` was generated from the rmarkdown file `ClimateMeta.Rmd` using `knitr`. Provided the data files are in the same directory as the rmarkdown file, the html document can be re-created. 
