---
title: "Climate data check"
author: "Aud Halbritter"
date: "18.01.2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
knitr::read_chunk("Climate Data Check.R")

library("ggplot2")
library("dplyr")
library("tidyr")
library("lubridate")

load("climate.Rdata")
load("climate_month.Rdata")

daily_dat <- climate %>% 
  mutate(day = ymd(format(dateTime, "%Y-%m-%d"))) %>%
  filter(variable %in% c("Tair", "Tsoil0", "Tsoil5", "Tsoil20")) %>% 
  group_by(day, site, variable, logger) %>% 
  summarise(mean = mean(value), min = min(value), max = max(value), diff = max - min) %>% 
  mutate(month = month(day))


```

# Data

Climate data at 4 sites, from the gradient and otc chambers (not all variables).

```{r, echo = FALSE}
table(climate$variable, climate$logger)
```


# Raw data

The worst spikes in the data that could be identified visually and duplicates were removed. The data are mostly recorded every ten minutes, but in the first file at each site, data are recorded every minute, at least for a while. I have left these values in the data.


# Checked variables

Tair: Air temperature in 째C at 2m for gradient, 30cm for OTC

Tsoil0: Ground temperature in 째C

Tsoil5: Soil temperature in 째C, 5cm below ground
Soil temperature not useful. For 3 sites only data in otc. For L site, the variation increases over time for the gradient (flagged)

Tsoil20: Soil temperature in 째C, 20cm below ground

waterContent5: Soil moisture in %, 5cm below ground
!!! NOT YET CHECKED !!!


## Raw data plotted

```{r HourlyDataTair, echo = FALSE}

```

```{r HourlyDataTGround, echo = FALSE}

```

```{r HourlyDataTsoil5, echo = FALSE}

```

```{r HourlyDataTsoil20, echo = FALSE}

```

## Have not removed spikes here! Need to figure out when the growing season starts.
```{r HourlyDatawaterContent5, echo = FALSE}

```




## Monthly Data

Air temperature at 2m along the gradient
```{r AirTemperatureGradient}

```


Ground and soil temperature data inside and outside of OTC
```{r TableWarmestMonth}

```



## Daily Data

### Daily difference between min and max values

```{r DailyDiffMinMax}

```
