---
title: "OTC climate"
author: "Richard J. Telford"
date: "September 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
load("otcc.Rdata")
load("otcc_month.Rdata")
load("month_annual2016-09-21.Rdata")

library("ggplot2")
library("dplyr")
```

## 

plots from otc


##Monthly OTC

```{r}
ggplot(otc_month, aes(x = month, y = value, colour = site)) + geom_path() + geom_point(size = 0.1) + facet_wrap(~variable, scales = "free_y")
```

##Comparison OTC with external weather

```{r}

all_month <- rbind(
  cbind(logger = "OTC", otc_month),
  cbind(logger = "Site", monthly)
)

g <- all_month %>% filter(variable == "Tsoil0") %>% 
  ggplot(aes(x = month, y = value, colour = logger)) + 
    geom_path() + 
    geom_point(size = 0.1) +
    facet_wrap(~site)

g

g %+% (all_month %>% filter(variable == "Tsoil5"))
g %+% (all_month %>% filter(variable == "Tsoil20"))
g %+% (all_month %>% filter(variable == "Tair"))
g %+% (all_month %>% filter(variable == "waterContent0"))
g %+% (all_month %>% filter(variable == "waterContent5"))
g %+% (all_month %>% filter(variable == "waterContent20"))
g %+% (all_month %>% filter(variable == "RH"))
