---
title: "Trait error checking"
author: "Richard J. Telford"
date: "November 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

#Some quick error checks on the 2016 trait data

```{r loadData}
#load libraries
library(ggplot2)
library(readr)
library(dplyr)

traits <- read_csv("data/2016_ChinaLeafTraitData_FlaggedErrors_11042016.csv")
problems(traits)#table entries that are incorrect mode for column (ie characters in numeric column). "Unknown" is actual entry in data.
#convert to numeric (has "too small")
traits$Leaf_Thickness_1_mm <- as.numeric(traits$Leaf_Thickness_1_mm)
```


## Leaf thicknesses
Some rather thick leaves.

```{r}
ggplot(traits, aes(x = Leaf_Thickness_1_mm, y = Leaf_Thickness_2_mm)) + 
  geom_point() +
  geom_abline(slope = 1, intercept = 0, colour = "red")
```

##Leaf thickness 3 against leaf thickness 1

Clear outliers
```{r}
ggplot(traits, aes(x = Leaf_Thickness_1_mm, y = Leaf_Thickness_3_mm)) + 
  geom_point() +  
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, colour = "red") #!!!
```

## Leaf dry mass against wet mass

Highly likely that points above green line are out by a factor of 1000 (i.e. mg rather than grams). Points between green and red lines (and slightly below this) are suspect.

```{r}
ggplot(traits, aes(x = Wet_Mass_g, y = Dry_Mass_g)) + 
  geom_point() + 
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, colour = "red") +#!!!
  geom_abline(slope = 1, intercept = 1, colour = "green") #!!!
```  
  
Correcting points where `log10(Dry_Mass_g) > log10(Wet_Mass_g) + 1` by dividing by 1000. Points that are still above the 1:1 (red) line are suspect, as are some of the small leaves and those below the cloud. NB increased variance expected with small leaves.
```{r}
traits %>%
  mutate(Dry_Mass_g = ifelse(log10(Dry_Mass_g) < log10(Wet_Mass_g) + 1, Dry_Mass_g, Dry_Mass_g/1000)) %>%
  ggplot(aes(x = Wet_Mass_g, y = Dry_Mass_g)) + geom_point() +
  scale_x_log10()+
  scale_y_log10() +
  geom_abline(slope = 1, intercept = 0, colour = "red") +#!!!
  geom_abline(slope = 1, intercept = 1, colour = "green") 
```

##Leaf wet mass vs area

Some clearly problematic leaves
```{r}
ggplot(traits, aes(x = Wet_Mass_g, y = Cropped_Leaf_Area)) + 
  geom_point() +  
  scale_x_log10() +
  scale_y_log10()
```

##Duplicated envelope names

Duplicate values in one (or both) of the raw data files in the `left_join`.

```{r duplicateNames}
traits %>% 
  group_by(Full_Envelope_Name) %>% 
  summarise(n = n()) %>% 
  filter(n > 1) %>%
  arrange(desc(n)) %>%
  print(n = 200)
```


